\section{Conclusion}
\label{sec:conclusion}
The Blockchain technology provides a way to find a total order of transactions
in a distributed system without relying on a trusted third party. This permits
to have an exact copy of the state in each node of the network, because each
node starts from the same state and changes the state according to the 
transactions, whose execution is deterministic. The main advantage of this 
technology with respect to traditional transition system, is its decentralized
nature.

In this report we took into consideration Ethereum, a representative of 
permissionless blockchain technology. This system is particularly interesting,
because it supports a general-purpose execution environment, the 
EVM (\autoref{sec:evm}).

In the first part of this work we propose a division of the Ethereum's 
architecture in logic layers and described each of them thoroughly in the 
respective sections.

The focus of the second part is the analysis of the scalability of Ethereum.
To do so, we analyzed the existing literature (\autoref{sec:background}) and 
confirmed empirically, that the current version of Ethereum reaches a maximal 
transaction throughput lower than two dozens per second, even if the
transactions do not require computations. This value is not influenced by the
number of miners, but rather by the block size and the block interval.
Furthermore, we analyzed the motivation of this reduced transaction throughput
with the aid of the cube of scalability`\cite{bib:art-of-scalability}. We 
argued that the current version of Ethereum is not developed in any direction 
of the cube and categorized the scalability improvement proposals, based on the 
axes they will affect. We concluded, that the main efforts of the community are 
concentrating on the z-axis of the cube~\autoref{sec:z-axis}, with proposals 
like Plasma and Sharding.

