\section{Conclusions}
\label{sec:conclusions}
The Blockchain technology provides a way to find a total order of transactions
in a distributed system without relying on a trusted third party. This allows
to have an exact copy of the state in each node of the network, because each
node starts from the same state and changes the state according to the
transactions, whose execution is deterministic. The main advantage of this
technology with respect to traditional transition systems is its decentralized
nature.

In this report we took into consideration Ethereum, a representative of the
permissionless blockchain technology. This system is particularly interesting
because it supports a general-purpose execution environment, the EVM.

In the first part of this work we proposed a decomposition of the Ethereum's
architecture in logical layers. The five stacked layers abstract from the
implementation details and offer a conceptual organization useful to compare
different blockchain proposals, although enabling flexibility as the case of the
EVM which is a vertical layer (see \autoref{sec:architecture}. Moreover, this
decomposition allows component-oriented development, which helps reasoning on the
system overall.

The focus of the second part is the analysis of the scalability of Ethereum. To
do so, we analyzed the existing literature (\autoref{sec:background}) and
confirmed empirically that the current version of Ethereum reaches a maximal
transaction throughput of approximately 15 transactions per second, even if the
transactions do not require computations. This value is not influenced by the
number of miners, but rather by the block size and the block interval.
Furthermore, we analyzed the motivation of this reduced transaction throughput
with the aid of the cube of scalability~\cite{bib:art-of-scalability}. We argue
that the current version of Ethereum is not developed in any direction of the
cube by analyzing the current specification. Thereafter, we categorize the
scalability improvement proposals based on the axes they will affect. From this
analysis, it is clear that the most efforts of the community are concentrated on
the z-axis of the cube (see \autoref{sec:z-axis}) with proposals like Plasma and
sharding. Sharding is a scaling strategy already seen in the database systems,
that is, systems which have to manage persistent data. Maybe, this affinity
between Ethereum and the database systems leads the z-axis of the Scale Cube.

The result of our research maps directly to the scalability
trilemma~\cite{bib:sharding-faq}, which states that we cannot have at the same
time scalability, decentralization and security. Assuming full security, we need
to find a trade-off between scalability and decentralization. The current
implementation of the Ethereum system provides security and decentralization and
gives up scalability.

Proposals like Proof-of-Stake (see~\autoref{sec:pos}) increase the transaction
throughput but it does not depend on the number of nodes in the system, hence,
we cannot view PoS as an improvement from the scalability point of view.
However, PoS increments the decentralization because validators do not require
powerful hardware.

Plasma does not change the protocol but provide a way to perform operations
off-chain

Sharding provides data partition allowing parallel processing transactions in
different shards, thus increasing the scalability of the system with respect to
the number of shards. However, the introduction of super nodes reduces the
decentralization of the system overall (see ~\autoref{sec:sharding-ethereum}).
