\subsection{Background}
\label{sec:background}

In this Section, we point out the usual concerns about the permissionless
blockchain system, in particular Bitcoin, and the related works regarding
Ethereum.

\paragraph{Bitcoin}
One major concern about Bitcoin and all permission-less cryptocurrencies in
general is the limited \textbf{maximum transaction
throughput}~\cite{bib:blockchain-challenges-opportunites-survey, bib:taxonomy,
bib:scaling-croman}. Thus, if the usage of this payment system augments, not all
transactions emitted can be processed in a predictable and bounded
time\footnote{\url{https://blockchain.info/charts/mempool-size}}. Moreover, the
choice of the transactions to include in the blockchain depends on the will of
the miners. Therefore, it is more likely that the more profitable transactions
are included in the distributed
ledger~\cite{bib:blockchain-challenges-opportunites-survey,wood2018ethereum}.

The maximum limited transaction throughput is due on the one hand to the $1$ MB
\emph{maximal block size} and on the other hand to the $10$ minutes \emph{block
interval}~\cite{bib:ghost, bib:scaling-croman}. Increasing the former would
certainly augment the transaction throughput and decrease the transaction fees,
but it would also slow down the propagation time, thus increasing the
possibility of forks. Decreasing the latter would increase the transaction
throughput and increase also the possibility of forks. The increase in the
number of forks reduces the security of the whole system~\cite{bib:ghost}. To
address these bottlenecks multiple solutions were proposed. The Blockchain
systems Bitcoin Cash and Litecoin try to augment the maximal transaction
throughput by modifying the Bitcoin client to have a maximum block size of $8$
MB and to have a block interval of $2.5$ minutes respectively. The
GHOST~\cite{bib:ghost} protocol consider also the stale blocks to permit to
diminish the block interval while preserving the same security guarantees of the
original proof of work as we already described in
\autoref{sec:consensus:algorithm}.

In addition to these considerations, a number of confirmation blocks are needed
to be sure that the transactions are really confirmed~\cite{bib:taxonomy}, i.e.
it is unlikely that the block in which they are included can be discarded in
favor of other blocks. This uncertainty in the acceptance of a block is also
known as \textbf{lack of Consensus Finality}~\cite{bib:the-quest} and is due to
the probabilistic nature of the PoW consensus algorithm. Currently, in Bitcoin
the number of confirmation blocks is set to $6$~\cite{bib:masteringbitcoin}.
This means that after the insertion of a transaction in a block at least an hour
must be waited to be sure that it would be confirmed.

Another concern with this technology is the need to store an increasingly large
amount of data to keep the desired security guarantees. Currently, the minimum
requirement to run a Bitcoin full-node, that is a node that verifies all the
transactions, is $145$
GB\footnote{\url{https://bitcoin.org/en/bitcoin-core/features/requirements}}.

Other major concerns are related to the \emph{cost} of the system, especially
the cost per confirmed transaction~\cite{bib:scaling-croman} which is due to the
mining cost, the transaction validation, the bandwidth and the storage.

Finally, the bootstrap time, that is the time needed for a new node to download
and process the whole transaction history, is another key factor that can
contribute to the scalability (scaling out)~\cite{bib:scaling-croman} and the
ability of new nodes to join the network.

\paragraph{Ethereum}
The previous considerations about the scalability of Bitcoin apply also to
Ethereum. In addition, the introduction of smart-contracts in Ethereum makes the
problem very similar to database replication and in particular \emph{state
machine replication}~\cite{bib:the-quest}. Also Wood~\cite{wood2018ethereum}
argues that it is very difficult to reach a high degree of scalability by
parallelizing transactions in this system because it is essentially a state
transaction machine. Indeed, the state in Ethereum influences the smart contract
execution and therefore the majority of transactions are dependent from previous
ones, thus making Ethereum \textbf{stateful}.

%*** et al. \cite{bib:sok:ethereum} individuate that scalability of blockchain
%is concerned with two major objectives: on the one hand increasing the
%transaction throughput and on the other hand decreasing the storage and
%bandwidth usage.
